---
title: "Hands-on_Ex03"
author: "Brandon Komajaya"
date: "30 April 2026"
date-modified: "last-modified"
format: html
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

Welcome to ISSS608, **VISUAL ANALYTICS AND APPLICATIONS**, this is a website containing my course work.

# Installing and loading the required libraries

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

# Reading Data File

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Interactivity

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = ID),
    stackgroups = TRUE, 
    binwidth = 1, 
    method = "histodot") +
  scale_y_continuous(NULL, 
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 6,
  height_svg = 6*0.618
)

```

# Interactivity (2)

```{r}
exam_data$tooltip <- c(paste0(     
  "Name = ", exam_data$ID,         
  "\n Class = ", exam_data$CLASS)) 

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(
    aes(tooltip = exam_data$tooltip), 
    stackgroups = TRUE,
    binwidth = 1,
    method = "histodot") +
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(
  ggobj = p,
  width_svg = 8,
  height_svg = 8*0.618
)
```

# Interactivity (3)

```{r}
tooltip_css <- "background-color:white; #<<
font-style:bold; color:black;" #<<

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(tooltip = ID),                   
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618,
  options = list(    #<<
    opts_tooltip(    #<<
      css = tooltip_css)) #<<
)                                        
```

# Alternative #1 : Separating by Race

```{r}

p <- ggplot(exam_data, aes(x = RACE, y = MATHS, fill = RACE)) +
  # 1. The Cloud (Density)
  # If you don't have ggdist installed, you can comment this out
  # stat_halfeye(adjust = .5, width = .6, justification = -.3, .width = 0, point_colour = NA) +
  
  # 2. The Rain (Dots)
  geom_dotplot_interactive(
    aes(tooltip = ID),
    binaxis = "y",         # Crucial: This tells R to bin along the MATHS scores
    stackdir = "center",   # Centers the dots on the race category
    method = "dotdensity", 
    dotsize = 0.8          # Start at 0.8 and decrease if they overlap
  ) +
  coord_flip() + 
  theme_minimal()

girafe(ggobj = p)
```

# Graphing Race

```{r}
tooltip <- function(y, ymax, accuracy = .01) {
  mean <- scales::number(y, accuracy = accuracy)
  sem <- scales::number(ymax - y, accuracy = accuracy)
  paste("Mean maths scores:", mean, "+/-", sem)
}

gg_point <- ggplot(data=exam_data, 
                   aes(x = RACE),
) +
  stat_summary(aes(y = MATHS, 
                   tooltip = after_stat(  
                     tooltip(y, ymax))),  
    fun.data = "mean_se", 
    geom = GeomInteractiveCol,  
    fill = "light blue"
  ) +
  stat_summary(aes(y = MATHS),
    fun.data = mean_se,
    geom = "errorbar", width = 0.2, size = 0.2
  )

girafe(ggobj = gg_point,
       width_svg = 8,
       height_svg = 8*0.618)
```

# Alternative #2: Added Mean and Standard Error

```{r}
library(tidyverse)
library(ggiraph)
library(scales) # for number formatting

# 1. Pre-calculate the mean, SE, and build the tooltip string using dplyr
exam_summary <- exam_data %>%
  group_by(RACE) %>%
  summarise(
    mean_maths = mean(MATHS, na.rm = TRUE),
    n = n(),
    sd_maths = sd(MATHS, na.rm = TRUE),
    # Standard Error (SE)
    se = sd_maths / sqrt(n),
    ymin = mean_maths - se,
    ymax = mean_maths + se,
    # Build a clean, formatted tooltip string
    tooltip = paste0(
      "<b>Mean Maths Score:</b> ", number(mean_maths, accuracy = 0.1),
      "<br><b>+/-:</b> ", number(se, accuracy = 0.1)
    )
  )

# 2. Build the chart using the pre-calculated data and native interactive geoms
# We map 'tooltip' in the main aes()
gg_clean <- ggplot(exam_summary, aes(x = RACE, y = mean_maths, tooltip = tooltip)) +
  # Use the specialized geom_col_interactive
  geom_col_interactive(fill = "lightblue", color = "white") +
  # Use geom_errorbar with the pre-calculated min/max
  geom_errorbar(aes(ymin = ymin, ymax = ymax), 
                width = 0.2, color = "black", size = 0.5) +
  # Cleaner styling
  theme_minimal() +
  labs(
    title = "Mean Maths Scores by Race with Standard Error",
    y = "Average Score",
    x = "Race/Ethnicity"
  )

# 3. Render the interactive chart
girafe(ggobj = gg_clean)
```

# Plotting Math

```{r}
p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(           
    aes(data_id = CLASS),             
    stackgroups = TRUE,               
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618                      
)                                        
```

# Plotting Math (Interactive)

```{r}
exam_data$onclick <- sprintf("window.open(\"%s%s\")",
"https://www.moe.gov.sg/schoolfinder?journey=Primary%20school",
as.character(exam_data$ID))

p <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(onclick = onclick),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +               
  scale_y_continuous(NULL,               
                     breaks = NULL)
girafe(                                  
  ggobj = p,                             
  width_svg = 6,                         
  height_svg = 6*0.618)                                        
```

# Maths and English Plotting

```{r}
p1 <- ggplot(data=exam_data, 
       aes(x = MATHS)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") +  
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

p2 <- ggplot(data=exam_data, 
       aes(x = ENGLISH)) +
  geom_dotplot_interactive(              
    aes(data_id = ID),              
    stackgroups = TRUE,                  
    binwidth = 1,                        
    method = "histodot") + 
  coord_cartesian(xlim=c(0,100)) + 
  scale_y_continuous(NULL,               
                     breaks = NULL)

girafe(code = print(p1 + p2), 
       width_svg = 6,
       height_svg = 3,
       options = list(
         opts_hover(css = "fill: #202020;"),
         opts_hover_inv(css = "opacity:0.2;")
         )
       ) 
```

# Different ways of displaying Graphic

```{r}
p1 <- plot_ly(data = exam_data, 
             x = ~MATHS, 
             y = ~ENGLISH)
p2 <- plot_ly(data = exam_data, 
        x = ~ENGLISH, 
        y = ~MATHS, 
        color = ~RACE)
p3 <- ggplot(data=exam_data, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

subplot(ggplotly(p1), ggplotly(p2),ggplotly(p3))
```

# Math / English and Math / Science

```{r}
d <- highlight_key(exam_data)
p1 <- ggplot(data=d, 
            aes(x = MATHS,
                y = ENGLISH)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))

p2 <- ggplot(data=d, 
            aes(x = MATHS,
                y = SCIENCE)) +
  geom_point(size=1) +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100))
subplot(ggplotly(p1),
        ggplotly(p2))
```

# Data Table

```{r}
DT::datatable(exam_data, class= "compact")

```

# Cross talk Data Table

```{r}
# 1. Create the shared data object
d <- highlight_key(exam_data)

# 2. Build the standard ggplot (Keep aes clean)
p <- ggplot(d, aes(ENGLISH, MATHS)) + 
  geom_point(size = 1) +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100))

# 3. Apply the highlighting here (This is where 'on' and 'off' live)
gg <- highlight(ggplotly(p), 
                on = "plotly_selected", 
                off = "plotly_deselect")

# 4. Display with crosstalk
crosstalk::bscols(gg, 
                  DT::datatable(d), 
                  widths = 5)   
```

# Alternative #3 :Added Filter to Interactive Table

```{r}
library(plotly)
library(crosstalk)
library(DT)
library(dplyr)

d <- SharedData$new(exam_data)

p <- ggplot(d, aes(ENGLISH, MATHS)) +
  geom_point(size = 2, alpha = 0.6) +
  theme_minimal() +
  coord_cartesian(xlim = c(0, 100), ylim = c(0, 100))


gg <- highlight(ggplotly(p), 
                on = "plotly_selected", 
                off = "plotly_deselect",
                persistent = FALSE)

bscols(
  widths = c(3, 9), # Left column width 3/12, Right column width 9/12
  list(
    filter_checkbox("gender", "Filter by Gender", d, ~GENDER),
    filter_slider("maths_score", "Maths Range", d, ~MATHS, width = "100%")
  ),
  list(
    gg,
    datatable(d, extensions = 'Scroller', options = list(scrollY = 200, scrollCollapse = TRUE))
  )
)
```

# Alternative #4 :Added Search Bar

```{r}
library(plotly)
library(crosstalk)
library(DT)

# 1. Initialize SharedData
d <- SharedData$new(exam_data)

# 2. Add Filter Widgets as a header row
# This allows users to narrow down data before selecting points
filter_row <- bscols(
  filter_select("race", "Select Race", d, ~RACE),
  filter_checkbox("gender", "Gender", d, ~GENDER, inline = TRUE)
)

# 3. Create the Plot
p <- ggplot(d, aes(ENGLISH, MATHS, color = RACE)) +
  geom_point(alpha = 0.5) +
  theme_minimal()

gg <- highlight(ggplotly(p), on = "plotly_selected", off = "plotly_deselect")

# 4. Display in a vertical stack
# widths = 12 makes them take up the full page width
bscols(widths = 12,
       filter_row,
       gg,
       datatable(d, options = list(pageLength = 5))
)
```

# Alternative #5 :Added Client Ready Table

```{r}
library(plotly)
library(crosstalk)
library(DT)

d <- SharedData$new(exam_data)

p <- ggplot(d, aes(ENGLISH, MATHS)) +
  geom_point(color = "steelblue", size = 1.5) +
  theme_light()

gg <- highlight(ggplotly(p), on = "plotly_selected", off = "plotly_deselect")

# Advanced DT with Buttons
# Note: 'dom = "Bfrtip"' is the magic string that enables buttons
custom_table <- datatable(
  d,
  extensions = 'Buttons',
  options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf'),
    pageLength = 10
  )
)

# Display Side-by-Side with specific width ratios
bscols(widths = c(6, 6),
       gg,
       custom_table
)
```

# Installing and loading the required libraries

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate, tidyverse)
```

```{r}
col <- c("Country", "Continent")
globalPop <- read_xls("data/GlobalPopulation.xls",
                      sheet="Data") %>%
  mutate(across(col, as.factor)) %>%
  mutate(Year = as.integer(Year))
```

# Plotting Static Population

```{r}
ggplot(globalPop, aes(x = Old, y = Young, 
                      size = Population, 
                      colour = Country)) +
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(title = 'Year: {frame_time}', 
       x = '% Aged', 
       y = '% Young') 
```

# Rendering Gif

```{r}
gg <- ggplot(globalPop, 
       aes(x = Old, 
           y = Young, 
           size = Population, 
           colour = Country)) +
  geom_point(aes(size = Population,
                 frame = Year),
             alpha = 0.7) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  labs(x = '% Aged', 
       y = '% Young') + 
  theme(legend.position='none')

ggplotly(gg)
```

# Animated Bubble Plot

```{r}
bp <- globalPop %>%
  plot_ly(x = ~Old, 
          y = ~Young, 
          size = ~Population, 
          color = ~Continent,
          sizes = c(2, 100),
          frame = ~Year, 
          text = ~Country, 
          hoverinfo = "text",
          type = 'scatter',
          mode = 'markers'
          ) %>%
  layout(showlegend = FALSE)
bp
```
